### 2025-09-26 YEAR2/AER210_LectureNotes/main.tex
sept 26

\begin{example}
    Consider the transformation and integral in \ref{ex:change_of_variables}. We can verify the result using the back transformation theorem. The Jacobian of the inverse transformation is:
    \begin{align*}
        J^{-1} &= \frac{\partial(u,v)}{\partial(x,y)} = \begin{vmatrix}
            \dfrac{\partial u}{\partial x} & \dfrac{\partial u}{\partial y} \\
            \dfrac{\partial v}{\partial x} & \dfrac{\partial v}{\partial y}
        \end{vmatrix} = \begin{vmatrix}
            2x & 2y \\
            2x & -2y
        \end{vmatrix} \\
        &= -4xy - 4xy = -8xy
    \end{align*}
    So, we have:
    $$        
        J = \frac{1}{J^{-1}} = -\frac{1}{8xy}
    $$
    This is consistent with the result we obtained in \ref{ex:change_of_variables}, and results in a much simpler integral to evaluate.
\end{example}
\chapter{Fluid Mechanics}
\end{document}

>>>>

\begin{example}
    Consider the transformation and integral in \ref{ex:change_of_variables}. We can verify the result using the back transformation theorem. The Jacobian of the inverse transformation is:
    \begin{align*}
        J^{-1} &= \frac{\partial(u,v)}{\partial(x,y)} = \begin{vmatrix}
            \dfrac{\partial u}{\partial x} & \dfrac{\partial u}{\partial y} \\
            \dfrac{\partial v}{\partial x} & \dfrac{\partial v}{\partial y}
        \end{vmatrix} = \begin{vmatrix}
            2x & 2y \\
            2x & -2y
        \end{vmatrix} \\
        &= -4xy - 4xy = -8xy
    \end{align*}
    So, we have:
    $$        
        J = \frac{1}{J^{-1}} = -\frac{1}{8xy} \implies I = \iint_R xy \, dA = \iint_S \frac{xy}{8xy} \, du \, dv = \iint_S \frac{1}{8} \, du \, dv = \frac{15}{8}
    $$
    This is consistent with the result we obtained in \ref{ex:change_of_variables}, and results in a much simpler integral to evaluate.
\end{example}

\begin{example}
    Consider the integral:  
    $$
        \iint_R \exp{\left(\frac{x - y}{x + y}\right)} \, dA
    $$
    where $R$ is the trapezoid with vertices at $(1,0)$, $(2,0)$, $(0,-2)$ and $(0,-1)$.

    Motivated by the integrabled, we can use the change of variables:
    $$
    \begin{cases}
        u = x + y \\
        v = x - y
    \end{cases} 
    $$
    Also, we can define our bounds as follow:
    \begin{itemize}
        \item $y = x - 2 \implies u = 2$
        \item $y = x - 1 \implies v = 1$
        \item $x = 0 \implies u = -v$
        \item $y = 0 \implies u = v$
    \end{itemize}
    We can find the Jacobian of this transformation by its inverse:
    $$
    J^{-1} = \frac{\partial(u,v)}{\partial(x,y)} = \begin{vmatrix}
        \dfrac{\partial u}{\partial x} & \dfrac{\partial u}{\partial y} \\
        \dfrac{\partial v}{\partial x} & \dfrac{\partial v}{\partial y}
    \end{vmatrix} = \begin{vmatrix}
        1 & 1 \\
        1 & -1
    \end{vmatrix} = -2 \implies J = -\frac{1}{2}
    $$
    So, we can write the integral in terms of $u$ and $v$:
    \begin{align*}
        \iint_R \exp{\left(\frac{x - y}{x + y}\right)} \, dA &= \iint_S \exp{\left(\frac{v}{u}\right)} \cdot |J| \, du \, dv \\
        &= \frac{1}{2} \int_1^2 \int_{-v}^{v} \exp{\left(\frac{v}{u}\right)} \, du \, dv \\
        &= \frac{1}{2} \int_1^2 v(e - e^{-1}) \, dv = \frac{3}{4} (e - e^{-1})
    \end{align*}
\end{example}

\begin{example}
    \textit{Evaluate the integral}
    $$
        \iint_R (x^2 - y^2) \exp(xy) \, dA
    $$
    where $R$ is the region bounded by the hyperbolas $xy=1$ and $xy=4$, and the lines $y=x$ and $y=x+2$.

    Motivated by the bounds, we can use the change of variables:
    $$
    \begin{cases}
        u = xy \\
        v = y - x
    \end{cases}
    $$
    So, let $S$ be the corresponding region in the $uv$-plane. We can find the bounds of $S$ by plugging in the lines that bound $R$:   
    \begin{itemize}
        \item $xy = 1 \implies u = 1$
        \item $xy = 4 \implies u = 4$
        \item $y = x \implies v = 0$
        \item $y = x + 2 \implies v = 2$
    \end{itemize}
    We find the Jacobian of this transformation by its inverse:
    $$
    J^{-1} = \frac{\partial(u,v)}{\partial(x,y)} = \begin{vmatrix}
        \dfrac{\partial u}{\partial x} & \dfrac{\partial u}{\partial y} \\
        \dfrac{\partial v}{\partial x} & \dfrac{\partial v}{\partial y}
    \end{vmatrix} = \begin{vmatrix}
        y & x \\
        -1 & 1
    \end{vmatrix} = y + x \implies J = \frac{1}{y + x}
    $$
    So, we can express the integral in terms of $u$ and $v$:
    \begin{align*}
        \iint_R (x^2 - y^2) \exp(xy) \, dA &= \iint_S (x+y)(x-y)e^u \cdot |J| \, du \, dv \\
        &= \int_0^2 \int_1^4 -v e^u \, du \, dv = \int_0^2 -v(e^4 - e) \, dv = 2(e - e^4)
    \end{align*}
\end{example}
\begin{example}
    \textit{Find the volume of the region bounded by the hyperbolic cylinders:}
    $$ 
    xy = 1, \quad xy = 3, \quad xz = 1, \quad xz = 3, \quad x = 36, \quad y = 25, \quad yz = 49
    $$
    Motivated by the bounds, we can use the change of variables:
    $$
    \begin{cases}
        u = xy \\
        v = xz \\
        w = yz
    \end{cases} 
    $$
    So, let $S$ be the corresponding region in the $uvw$-space. We can find the bounds of $S$ by plugging in the surfaces that bound the region:

    \begin{align*}
        1 &\leq u \leq 3 \\
        1 &\leq v \leq 3 \\
        25 &\leq w \leq 49
    \end{align*}

\end{example}
\chapter{Fluid Mechanics}
\end{document}

### 2025-09-26 YEAR2/ECE253_LectureNotes/main.tex
sept 26

\begin{definition}[Gated D-Latch]
    We can set $D = S = \overline{R}$, so we would prevent $S = R = 1$. So when the clock is HI, $Q$ tracks $D$ (a.k.a. Transparent Mode), and when clock is high, it won't store (a.k.a. Opaque Mode). The truth table is shown as follows:
    \begin{table}[h!]
        \centering
        \begin{tabular}{|c|c||c|}
            \hline
            CLK & $D$ & $Q(t+1)$ \\
            \hline
            0 & X & $Q(t)$ \\
            1 & 0 & 0 \\
            1 & 1 & 1 \\ 
            \hline
        \end{tabular}
        \caption{Truth Table for Gated D Latch}
        \label{tab:gated_d_latch_truth_table}
    \end{table}
\end{definition}

@@

\begin{definition}[D-Flip Flops]
    A D flip-flop is a type of flip-flop that captures the value of the D input at a specific portion of the clock cycle (rising edge) and holds that value until the next clock cycle. It is done by chaining the output of a D latch to the input of another D-latch. While the second D-latch recognizes the CLK, and the first one recognizes $\overline{\text{CLK}}$.
\end{definition}
\section{Finite State Machines (FSM)}
% FSM modeling, state diagrams, and Verilog implementation (to be expanded)

>>>>

\begin{definition}[Gated D-Latch]
    We can set $D = S = \overline{R}$, so we would prevent $S = R = 1$. So when the clock is HI, $Q$ tracks $D$ (a.k.a. Transparent Mode), and when clock is high, it won't store (a.k.a. Opaque Mode). The truth table is shown as follows:
    \begin{table}[h!]
        \centering
        \begin{tabular}{|c|c||c|}
            \hline
            CLK & $D$ & $Q(t+1)$ \\
            \hline
            0 & X & $Q(t)$ \\
            1 & 0 & 0 \\
            1 & 1 & 1 \\ 
            \hline
        \end{tabular}
        \caption{Truth Table for Gated D Latch}
        \label{tab:gated_d_latch_truth_table}
    \end{table}

    The Verilog implementation is as follows:
    \begin{verbatim}
    module d_latch(input logic D, CLK,
                   output logic Q);
        always_latch begin // this is new in SystemVerilog, 
        // dont write else in always_latch 
            if (CLK) // Transparent mode
                Q = D;
        end
    endmodule
    \end{verbatim}
\end{definition}

@@

\begin{definition}[D-Flip Flops (Registors)]
    A D flip-flop is a type of flip-flop that captures the value of the D input at a specific portion of the clock cycle (rising edge) and holds that value until the next clock cycle. It is done by chaining the output of a D latch to the input of another D-latch. While the second D-latch recognizes the CLK, and the first one recognizes $\overline{\text{CLK}}$.

    The verilog implementation is as follows:
    \begin{verbatim}
    module d_ff(input logic D, CLK,
                output logic Q);
        always_ff @(posedge CLK) begin
            Q <= D; // Non-blocking assignment
        end
    endmodule
    \end{verbatim}
    \verb|postedge| is a keyword used to create a positive edge-triggered FF. We use $<= $ instead of $= $ because we want to use non-blocking assignment. This is because in a sequential circuit, all the FFs are triggered at the same time, and we want to make sure that the value of Q is updated only after all the FFs have been evaluated.
\end{definition}

\begin{definition}[8-Bit Register]
    An 8-bit register is a digital storage device that can hold 8 bits of binary data. It is typically implemented using a series of D flip-flops, where each flip-flop stores one bit of data. The register has a clock input (CLK) that controls when the data is loaded into the register, and an enable input (EN) that determines whether the register should load new data or retain its current value. When EN is high (1), the register loads the data present on the D input into the flip-flops on the rising edge of the clock. When EN is low (0), the register retains its current value regardless of changes on the D input.

    The Verilog implementation is as follows:
    \begin{verbatim}
    module register8(input logic [7:0] D,
                     input logic CLK, EN,
                     output logic [7:0] Q);
        always_ff @(posedge CLK) begin
            if (EN)
                Q <= D; // Load new data
            // else retain old data
        end
    endmodule
    \end{verbatim}

\paragraph{Resets} We have two types of rests to reset the FF to a known state:
\begin{itemize}
    \item \textbf{Synchronous reset:} Active only on the clock edge. If reset = 1 at the triggering edge the flip‑flop is forced to the predefined state (usually 0); otherwise it captures D.
    \item \textbf{Asynchronous reset:} Active immediately when asserted, forcing the flip‑flop to the predefined state (usually 0) regardless of the clock; normal edge‑triggered operation resumes when deasserted.
\end{itemize}
\begin{definition}[Synchronous Reset D-FF]
    The Verilog implementation is as follows:
    \begin{verbatim}
    module d_ff_sync_reset(input logic D, CLK, RST,
                           output logic Q);
        always_ff @(posedge CLK) begin
            if (RST==1'b0)
                Q <= 1'b0; // Reset to 0
            else
                Q <= D; // Normal operation
        end
    endmodule
    \end{verbatim}

\end{definition}
\section{Finite State Machines (FSM)}
% FSM modeling, state diagrams, and Verilog implementation (to be expanded)
